Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-06-11 Paper/paper.tex
Update text
Gabor's input.

\begin{abstract}
  Determinations of the UV luminosity function of AGN at high
  redshifts are important for constraining the AGN contribution to
  reionization and understanding the growth of supermassive black
  holes.  Recent inferences of the luminosity function suffer from
  inconsistencies arising from inhomogeneous selection and mixing of
  AGN data.  We address this problem by constructing a sample of more
  than 80,000 colour-selected AGN from redshift $z=0$ to $7.5$.  While
  this sample is composed of multiple data sets with spectroscopic
  redshifts and completeness estimates, we homogenise these data sets
  to identical cosmologies, intrinsic AGN spectra, and magnitude
  systems.  Using this sample, we derive the AGN UV luminosity
  function from redshifts $z=0$ to $7.5$.  The luminosity function has
  a double power law form at $z<5.5$, and a single power law form at
  higher redshifts.  The break luminosity brightens rapidly with
  redshift and the faint end of the luminosity function steepens.  In
  spite of this steepening, the contribution of AGN to the hydrogen
  photoionization rate at $z\sim 6$ is subdominant, although it can be
  non-negligible if these luminosity functions hold down to
  $M_{1450}=-18$.  Under reasonable assumptions, AGN can reionize
  \HeII\ by redshift $z=3.2$.  At low redshifts ($z<0.5$), AGN can
  produce all of the hydrogen photoionization rate inferred from the
  statistics of \HI\ absorption lines in the IGM.  Our global analysis
  of the luminosity function also reveals important systematic errors
  in the data, particularly at $z=2.2$--$3.5$, which need to be
  addressed in future in order to improve our results.  We make
  various fitting functions, luminosity function analysis codes, and
  homogenized AGN data publicly available.
\end{abstract}

@@

-- some people use photo-zs
-- some people mix binned data
-- some people fix slopes
-- some people exclude some data arbitrarily

@@

\subsection{Sample Selection}

We started with
%% JFH we started with --> we started by
compiling the samples of recent photometric rest-frame UV-optical
quasar surveys. The restriction to UV-optical surveys was mainly
driven by our science goal to characterise the UV luminosity function
of Type~1 quasars. X-ray-selected samples are less suited for this
purpose due to spectroscopic incompleteness and the $\sim 0.4$\,dex
scatter in the conversion from X-ray to UV luminosity
\citep{2010A&A...512A..34L, 2015MNRAS.453.1946G, 2016ApJ...819..154L}
that contributes significantly to the error budget in the UV
luminosity function of X-ray-selected samples unless rest-frame UV
photometry is incorporated \citep{2015AA...578A..83G}.  The individual
surveys and their main characteristics are listed in
Table~\ref{tab:samples}.  Figure~\ref{fig:qsos} presents a redshift
histogram of the contributing surveys.

We included surveys based on a set of simple criteria:
\begin{enumerate}
\item Spectroscopic redshifts for the vast majority of targets.
  %% JFH Did we actually use any photometric redshifts? If not, I would omit this ``Vast majority'' and make
  %% it more definite. 
\item Accurate rest-frame UV-optical CCD photometry.
\item Statistical power (sample size, coverage in $z$ and/or absolute magnitude).
\item A characterised selection function.
  %% JFH well characterized
\end{enumerate}
As a prerequisite for a joint analysis of the QLF we obtained the
survey selection functions in electronic form, either from the
publication or
%% JFH by request 
from the authors. As a reference for future surveys we
publish them here
%% JFH rather than publish them here say ``we make them electronically available here''. Cite an appendix where
%% you do this. 
in modified and homogenised form
(Section~\ref{sect:datahom}).

Due to their selection criteria and their statistical power specific
surveys contribute to distinct redshift ranges. At $z<2.2$ we
considered quasars from the SDSS DR7 quasar catalogue
\citep{2010AJ....139.2360S} and the 2SLAQ survey catalogue
\citep{2009MNRAS.392...19C}. We restricted the SDSS DR7 sample to the
48,664 $0.1<z<2.2$ quasars selected with the final SDSS quasar
selection algorithm \citep{2002AJ....123.2945R, 2006AJ....131.2766R}
from a survey area of 6248\,deg$^2$ \citep{2012ApJ...746..169S}. We
adopted the SDSS targeting photometry corrected for Galactic
extinction \citep{2010AJ....139.2360S}. To limit systematic
uncertainties in the correction for host galaxy light
%% JFH What is Croom the reference for here, host galaxy contamination or the 2SLAQ sample. You basically
%% need a reference for 2SLAQ and it appears in the wrong place here. 
\citep{2009MNRAS.392...19C} we restricted the 2SLAQ sample to 11,816
$g<21.85$ $0.4<z<2.2$ quasars from its spectroscopic survey footprint
near the North Galactic Pole (NGP, 8153 quasars in $127.7$\,deg$^2$)
and the South Galactic Pole (SGP, 3663 quasars in
$64.2$\,deg$^2$). The small sample overlap between SDSS and 2SLAQ (112
quasars) has negligible impact on the QLF evaluation.
%% JFH Maybe state what the 2SLAQ survey goal was re quasars.

@@

At $4.7\le z<5.5$ we combined several recent surveys, accounting for
sample overlap and updated selection functions. At the bright end of
the QLF we used the 99 quasars from the SDSS+WISE survey
\citep{2016ApJ...829...33Y} that have $M_{1450}<-26.73$ in our adopted
cosmology.  For these 99 quasars selected from 14,555\,deg$^2$ we
adopted the \citet{2016ApJ...829...33Y} selection function.  The
\citet{2016ApJ...829...33Y} sample partially overlaps with the SDSS
DR7 sample from \citet{2013ApJ...768..105M}, so to avoid
double-counting quasars we used the latter sample only at $M_{1450}\ge
-26.73$, yielding 104 additional $4.7\le z<5.5$ quasars selected in
6248\,deg$^2$. We used the $z\sim 5$ SDSS DR7 selection function from
\citet{2013ApJ...768..105M} that supersedes the one from
\citet{2006AJ....131.2766R} due to improved bandpass corrections and IGM
parameterization.  To these two bright-end samples we added the
faint-end sample from the \citet{2013ApJ...768..105M} SDSS Stripe~82
survey (62 uniformly selected $4.7\le z<5.5$ quasars in 235\,deg$^2$)
and two $4.7\le z<5.5$ quasars from \citet{2011ApJ...728L..26G},
adopting the respective selection functions.  We did not consider the
limit on the $z\sim 5$ QLF by \citet{2012ApJ...756..160I} due to
systematic errors in their selection
function\footnote{\citet{2012ApJ...756..160I} underestimated the
  dispersion in rest-frame UV quasar colours with respect to SDSS at
  all redshifts (their Figure~4). Contrary to their claim, photometric
  errors have a small effect on the colour distribution of SDSS
  quasars given the statistical errors of $<0.03$\,mag in $gri$ for 90
  per cent of the SDSS DR7 bright quasar sample ($i<19.1$) and a
  relative calibration error of $\sim 1$ per cent
  \citep{2008ApJ...674.1217P}.}.

The SDSS colours of $5.1<z<5.5$ quasars are similar to those of M and
L dwarf stars, resulting in a low and uncertain selection efficiency
\citep{2013ApJ...768..105M}. WISE mid-infrared selection performs
better \citep{2016ApJ...829...33Y}, but is restricted to the bright
end of the quasar population. Unlike \citet{2013ApJ...768..105M} we
include the 9 uniformly selected $M_{1450}\ge -26.73$ SDSS DR7 quasars
and the 10 SDSS Stripe~82 quasars at $z>5.1$, adopting their low
selection efficiency.
%% JFH efficiency and completeness are two different things and you seem to be conflating
%% the two concepts here. 
As we will show in Section~3.2, the resulting
QLF is consistent with those at lower and higher redshifts, indicating
that the \citet{2013ApJ...768..105M} selection functions are quite
reliable.

@@

At the highest redshifts $z>6.5$ we considered two quasars found by
\citet{2011Natur.474..616M} and \citet{2015ApJ...801L..11V} in UKIDSS
imaging (3370\,deg$^2$)
%% JFH quote the individual redshifst of the mortlock and venemans objects
, in addition to the current highest-redshift
quasar at $z=7.54$ selected from a combination of UKIDSS, WISE and
DECaLS \citep[$\sim 2500$\,deg$^2$, ][]{2018Natur.553..473B}. Although
currently only rough estimates exist concerning their selection
functions, these quasars provide constraints on the evolution of the
integrated quasar space density from $z\sim 6$ to $z\sim 7$.  We use
them in our secondary analysis in Section~\ref{sec:global}.  Although
we do not include the highly debated \citet{2015AA...578A..83G} sample
in our main analysis due to its rough selection function and lacking
%% JFH lacking --> lack of 
spectroscopy for 17 of the 22 quasar candidates, we use it in
Section~\ref{sec:global} to constrain the faint end ($M_{1450}>-23$)
of the QLF at $z>4.1$. We restricted the \citet{2015AA...578A..83G}
sample to the 19/22 sources considered in their QLF.

@@

%% JFH We need a better roadmap here for where this section is going. 1) Binned estimates 2) individual
%% fits in redshift bins. 3) examining redshift evolution. 4) attempting a joint fit.

After homogenising the samples and selection functions we are in a position
to compute the UV luminosity function of AGN. We begin our analysis by
deriving luminosity functions in bins of redshift and absolute magnitude.
For simplicity we will use the notation $M\equiv M_{1450}$ in the following.

@@

In a magnitude bin $[M_\mathrm{min}, M_\mathrm{max})$, and redshift
  bin $[z_\mathrm{min}, z_\mathrm{max})$, we define the luminosity
    %% JFH Notation error here. 
function as \citep{2000MNRAS.311..433P}
\begin{equation}
  \phi \equiv \frac{N}{V_\mathrm{bin}},
\end{equation}
where $N$ is the number of quasars with magnitude
$M_\mathrm{min}\leq M<M_\mathrm{max}$ and redshift
$z_\mathrm{min}\leq z<z_\mathrm{max}$, and
\begin{equation}
  V_\mathrm{bin} = \int_{M_\mathrm{min}}^{M_\mathrm{max}}\mathrm{d}M
  \int_{z_\mathrm{min}}^{z_\mathrm{max}}\mathrm{d}z\, f(M, z)\,\frac{\mathrm{d}V}{\mathrm{d}z},
  \label{eqn:vi}
\end{equation}
is the effective volume of the bin. The inclusion of the survey selection function
$f(M,z)$ (Section~\ref{sect:datahom}) in Equation~(\ref{eqn:vi}) accounts for
what are sometimes called ``incomplete bins''
\citep{2006AJ....131.2766R}.  The comoving volume element $\mathrm{d}V/\mathrm{d}z$ is
given by
\begin{equation}
  \frac{\mathrm{d}V}{\mathrm{d}z}=\frac{\mathrm{d}V}{\mathrm{d}z\,\mathrm{d}\Omega}\times A\times\frac{4\pi}{41253},
\end{equation}
where $A$ is the survey area in deg$^2$, and 
\begin{equation}
  \frac{\mathrm{d}V}{\mathrm{d}z\,\mathrm{d}\Omega}=\frac{c}{H_0}\frac{d_L^2\left(z\right)}
       {\left(1+z\right)^2\sqrt{\Omega_\mathrm{m}\left(1+z\right)^3+\Omega_\Lambda}}
  \label{eqn:dvdzdo}
\end{equation}
denotes the comoving volume element per unit solid angle
\citep{1999astro.ph..5116H}.

The luminosity function $\phi$ has units of
$\mathrm{cMpc}^{-3}\mathrm{mag}^{-1}$.  We evaluate the double
integral in Equation~(\ref{eqn:vi}) by the Euler method, i.e., by
simply summing over the ``tiles'' of the selection function map,
without interpolating between the redshift and luminosity values of
neighbouring tiles. 
%%GW: We should add that interpolation of selection functions is not
%%straightforward (strong gradients) and not unique. Many selection
%%functions are very coarse. V=0 means that the selection function has
%%systematic errors (a QSO was found where the selection probability
%%is zero). By removing them we avoid numerical issues, but
%%underestimate the QLF. Surveys need to report selection functions
%%that are consistent with their actual samples. -- done (GK) 
This may result in $V_i=0$ for some quasars, in which case we remove
them from our analysis.  (Interpolation of selection functions is not
straightforward due to strong gradients.  The presence of objects with
$V_i=0$ implies that the selection function has systematic errors.)
%%GW: Poisson statistics will underestimate the error since
%%uncertainties in the selection function are not considered (all
%%surveys assume a fixed selection function). For small surveys the
%%Poisson error dominates, but for large surveys the error is driven
%%by systematics in the selection function. These facts have been
%%neglected so far, and we should make this point very clearly. -- done (GK)
In each bin, we estimate the uncertainty in the luminosity function by
assuming Poisson statistics \citep{1986ApJ...303..336G} for the number
of quasars, i.e.\ assuming negligible uncertainty in the selection function.
The resultant binned luminosity function estimates are shown by the
points in Figure~\ref{fig:mosaic}.

%%GW: I've sharpened the discussion below. -- done (GK)
As seen in Figure~\ref{fig:mosaic}, the distribution of luminosity
function values in each redshift bin are suggestive of a double power
law form for the QLF.  We will fit such a form below. 
However, in several redshift bins we note a suspicious decline of the
luminosity function at the faint limit of several surveys, for example in
the 2SLAQ sample at $z<2.2$, and in the SDSS sample at $z<1.8$ and $z\sim 4$.
The inconsistency between the SDSS faint end and the deeper
2SLAQ QLF indicates that the SDSS selection function is
systematically overestimated at its magnitude limit.
At $z\sim 4$ the SDSS faint end QLF is inconsistent with the fainter
\citet{2011ApJ...728L..26G} QLF. We identify such discrepant bins
by eye and discard them from our analysis.
%%GW: We should also set the selection functions to zero in these
%%bins. -- done (GK)
The discarded magnitude bins are shown in Figure~\ref{fig:mosaic} by
open circles.

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{mosaic_small.pdf}
    % mosaic.py 
  \end{center}
  \caption{Homogenised quasar luminosity functions at
    $\lambda=1450$\,\AA\ in redshift bins from $z=0.1$ to $6.5$.  The
    symbols show our inferred binned luminosity functions from various
    data sets:
    \citet[red]{2010AJ....139.2360S},
    \citet[green]{2009MNRAS.392...19C},
    \citet[dark blue]{2013ApJ...773...14R},
    \citet[light blue]{2011ApJ...728L..26G},
    \citet[yellow]{2016ApJ...829...33Y},
    \citet[DR7 brown]{2013ApJ...768..105M},
    \citet[Stripe 82 teal]{2013ApJ...768..105M},
    \citet[pink]{2016ApJ...833..222J},
    \citet[orange]{2010AJ....139..906W}, and
    \citet[grey]{2015ApJ...798...28K}.
    Open circles in corresponding colours indicate magnitude bins
    excluded due to incompleteness for the respective data sets.  The
    number of AGN before this selection is shown in parentheses; the
    selected number of AGN is shown outside parentheses.  In each
    redshift bin, the black curve shows our fiducial double power law
    model fit, which is represented by the median of the posterior
    probability distribution function.  The grey shaded area shows the
    one-sigma (68.26\%) uncertainty.  See Sections~\ref{sec:binnedlf}
    and \ref{sec:bins} for further details.}
  \label{fig:mosaic}
\end{figure*}

@@

%%GW: I would delete the z_bin column from the table. Otherwise the
%%reader might be confused what you are plotting in the figure. The
%%bin centers and the mean values are very similar anyway.--done (GK)
%%GW: Again 2.5<z<2.6 is missing--done (GK)
%%GW: I would add a column or a comment which redshift bins we
%%consider reliable. I'm afraid lazy people will just take the values
%%without thinking.
\begin{table*}
  % bins_tabulate.py; Nqso added by hand
  \caption{Derived double power law luminosity function parameters in
    various redshift bins.  The luminosity function parameters
    $\phi_*$, $M_*$, $\alpha$, and $\beta$ are defined in
    Equation~(\ref{eqn:dpl}), with $\beta$ denoting the faint-end
    slope.  Quasars in each bin have redshifts $z_\mathrm{min}\leq z <
    z_\mathrm{max}$.  The number of quasars in each bin is given by
    $N_\mathrm{qso}$.  The bin centre and sample mean redshift are
    given by $z_\mathrm{bin}$ and $\langle z\rangle$, respectively.
    Errors indicate one-sigma (68.26\%) uncertainties.  These values
    are shown in Figure~\ref{fig:evoln}.  The corresponding luminosity
    functions are shown in Figure~\ref{fig:mosaic}.  See
    Section~\ref{sec:bins} for further details.}
  \label{tab:bins}
  \begin{tabular}{cccr....}
    \hline
    $\langle z\rangle$ &
    $z_\mathrm{min}$ &
    $z_\mathrm{max}$ &
    $N_\mathrm{qso}$ &
    \multicolumn{1}{c}{$\log_{10}(\phi_*/$} &
    \multicolumn{1}{c}{$M_*$} &
    \multicolumn{1}{c}{$\alpha$} &
    \multicolumn{1}{c}{$\beta$} \\

>>>>

\begin{abstract}
  Determinations of the UV luminosity function of AGN at high
  redshifts are important for constraining the AGN contribution to
  reionization and understanding the growth of supermassive black
  holes.  Recent inferences of the luminosity function suffer from
  inconsistencies arising from inhomogeneous selection and mixing of
  AGN data.  We address this problem by constructing a sample of more
  than 80,000 colour-selected AGN from redshift $z=0$ to $7.5$.  While
  this sample is composed of multiple data sets with spectroscopic
  redshifts and completeness estimates, we homogenise these data sets
  to identical cosmologies, intrinsic AGN spectra, and magnitude
  systems.  Using this sample, we derive the AGN UV luminosity
  function from redshifts $z=0$ to $7.5$.  The luminosity function has
  a double power law form at $z<5.5$, and a single power law form at
  higher redshifts.  The break luminosity brightens rapidly with
  redshift and the faint end of the luminosity function steepens.  In
  spite of this steepening, the contribution of AGN to the hydrogen
  photoionization rate at $z\sim 6$ is subdominant, although it can be
  non-negligible if these luminosity functions hold down to
  $M_{1450}=-18$.  Under reasonable assumptions, AGN can reionize
  \HeII\ by redshift $z=3.2$.  At low redshifts ($z<0.5$), AGN can
  produce all of the hydrogen photoionization rate inferred from the
  statistics of \HI\ absorption lines in the IGM.  Our global analysis
  of the luminosity function also reveals important systematic errors
  in the data, particularly at $z=2.2$--$3.5$, which need to be
  addressed in future in order to improve our results.  We make
  various fitting functions, luminosity function analysis codes, and
  homogenised AGN data publicly available.
\end{abstract}

@@

%% -- some people use photo-zs
%% -- some people mix binned data
%% -- some people fix slopes
%% -- some people exclude some data arbitrarily

@@

\subsection{Sample Selection}
\label{sect:samplesel}

We started by compiling the samples of recent photometric rest-frame UV-optical
quasar surveys. The restriction to UV-optical surveys was mainly
driven by our science goal to characterise the UV luminosity function
of Type~1 quasars. X-ray-selected samples are less suited for this
purpose due to spectroscopic incompleteness and the $\sim 0.4$\,dex
scatter in the conversion from X-ray to UV luminosity
\citep{2010A&A...512A..34L, 2015MNRAS.453.1946G, 2016ApJ...819..154L}
that contributes significantly to the error budget in the UV
luminosity function of X-ray-selected samples unless rest-frame UV
photometry is incorporated \citep{2015AA...578A..83G}.  The individual
surveys and their main characteristics are listed in
Table~\ref{tab:samples}.  Figure~\ref{fig:qsos} presents a redshift
histogram of the contributing surveys.

We included surveys based on a set of simple criteria:
\begin{enumerate}
\item High spectroscopic completeness of the target sample.
\item Accurate rest-frame UV-optical CCD photometry.
\item Statistical power (sample size, coverage in $z$ and/or absolute magnitude).
\item A well-characterised selection function.
\end{enumerate}
As a prerequisite for a joint analysis of the QLF we obtained the
survey selection functions in electronic form, either from the
publication or by request from the authors. As a reference for future surveys we
make them electronically available here
%% JFH rather than publish them here say ``we make them electronically available here''. Cite an appendix where
%% you do this. 
in modified and homogenised form
(Section~\ref{sect:datahom}).

Due to their selection criteria and their statistical power specific
surveys contribute to distinct redshift ranges. At $z<2.2$ we
considered quasars from the SDSS DR7 quasar catalogue
\citep{2010AJ....139.2360S} and the 2SLAQ survey catalogue
\citep{2009MNRAS.392...19C}. We restricted the SDSS DR7 sample to the
48,664 $0.1<z<2.2$ quasars selected with the final SDSS quasar
selection algorithm \citep{2002AJ....123.2945R, 2006AJ....131.2766R}
from a survey area of 6248\,deg$^2$ \citep{2012ApJ...746..169S}. We
adopted the SDSS targeting photometry corrected for Galactic
extinction \citep{2010AJ....139.2360S}. To limit systematic
uncertainties in the correction for host galaxy light
\citep[detailed in][]{2009MNRAS.392...19C} we restricted the 2SLAQ sample to 9365
$g<21.85$ $0.4<z<2.2$ quasars from its spectroscopic survey footprint
near the North Galactic Pole (NGP, 7027 quasars in $127.7$\,deg$^2$)
and the South Galactic Pole (SGP, 2338 quasars in
$64.2$\,deg$^2$). The small sample overlap between SDSS and 2SLAQ (102
quasars) has negligible impact on the QLF evaluation.

@@

At $4.7\le z<5.5$ we combined several recent surveys, accounting for
sample overlap and updated selection functions. At the bright end of
the QLF we used the 99 quasars from the SDSS+WISE survey
\citep{2016ApJ...829...33Y} that have $M_{1450}\le -26.73$ in our adopted
cosmology.  For these 99 quasars selected from 14,555\,deg$^2$ we
adopted the \citet{2016ApJ...829...33Y} selection function.  The
\citet{2016ApJ...829...33Y} sample partially overlaps with the SDSS
DR7 sample from \citet{2013ApJ...768..105M}, so to avoid
double-counting quasars we used the latter sample only at
$M_{1450}>-26.73$, yielding 103 additional $4.7\le z<5.5$ quasars selected in
6248\,deg$^2$. We used the $z\sim 5$ SDSS DR7 selection function from
\citet{2013ApJ...768..105M} that supersedes the one from
\citet{2006AJ....131.2766R} due to improved bandpass corrections and IGM
parameterization.  To these two bright-end samples we added the
faint-end sample from the \citet{2013ApJ...768..105M} SDSS Stripe~82
survey (59 uniformly selected $M_{1450}>-26.73$ $4.7\le z<5.5$ quasars in 235\,deg$^2$)
and two $4.7\le z<5.5$ quasars from \citet{2011ApJ...728L..26G},
adopting the respective selection functions.  We did not consider the
limit on the $z\sim 5$ QLF by \citet{2012ApJ...756..160I} due to
systematic errors in their selection
function\footnote{\citet{2012ApJ...756..160I} underestimated the
  dispersion in rest-frame UV quasar colours with respect to SDSS at
  all redshifts (their Figure~4). Contrary to their claim, photometric
  errors have a small effect on the colour distribution of SDSS
  quasars given the statistical errors of $<0.03$\,mag in $gri$ for 90
  per cent of the SDSS DR7 bright quasar sample ($i<19.1$) and a
  relative calibration error of $\sim 1$ per cent
  \citep{2008ApJ...674.1217P}.}.

The SDSS colours of $5.1<z<5.5$ quasars are similar to those of M and
L dwarf stars, resulting in a low and uncertain completeness
\citep{2013ApJ...768..105M}. WISE mid-infrared selection performs
better \citep{2016ApJ...829...33Y}, but is restricted to the bright
end of the quasar population. Unlike \citet{2013ApJ...768..105M} we
include the 9 uniformly selected $M_{1450}>-26.73$ SDSS DR7 quasars
and the 10 SDSS Stripe~82 quasars at $z>5.1$, adopting their low
completeness. As we will show in Section~3.2, the resulting
QLF is consistent with those at lower and higher redshifts, indicating
that the \citet{2013ApJ...768..105M} selection functions are quite
reliable.

@@

At the highest redshifts $z>6.5$ we considered ULAS~J1120$+$0641
\citep[$z=7.085$,][]{2011Natur.474..616M} and PSO~J$036.5078+03.0498$
\citep[$z=6.527$,][]{2015ApJ...801L..11V}, both discovered in UKIDSS imaging
(3370\,deg$^2$), in addition to the current highest-redshift quasar J1342$+$0928
at $z=7.54$ selected from a combination of UKIDSS, WISE and DECaLS
\citep[$\sim 2500$\,deg$^2$,][]{2018Natur.553..473B}. Although
currently only rough estimates exist concerning their selection
functions, these quasars provide constraints on the evolution of the
integrated quasar space density from $z\sim 6$ to $z\sim 7$.  We use
them in our secondary analysis in Section~\ref{sec:global}.  Although
we do not include the highly debated \citet{2015AA...578A..83G} sample
in our main analysis due to its rough selection function and lack of
spectroscopy for 17 of the 22 quasar candidates, we use it in
Section~\ref{sec:global} to constrain the faint end ($M_{1450}>-23$)
of the QLF at $z>4.1$. We restricted the \citet{2015AA...578A..83G}
sample to the 19/22 sources considered in their QLF.

@@

%% JFH We need a better roadmap here for where this section is going. 1) Binned estimates 2) individual
%% fits in redshift bins. 3) examining redshift evolution. 4) attempting a joint fit.
%% GW: Agreed. Modified the roadmap.

After homogenising the samples and selection functions we are in a position
to compute the UV luminosity function of AGN. We begin our analysis by computing
binned estimates of the luminosity function as a function of magnitude in several redshift intervals.
We then perform parametric maximum-likelihood fits of the luminosity function in the individual redshift bins,
examine the resulting parameters as a function of redshift, and attempt a joint fit in magnitude and redshift.
For simplicity we will use the notation $M\equiv M_{1450}$ in the following.

@@

In a magnitude bin $[M_\mathrm{min}, M_\mathrm{max})$, and redshift
  bin $[z_\mathrm{min}, z_\mathrm{max})$, we define the binned luminosity
    %% JFH Notation error here. 
function as \citep{2000MNRAS.311..433P}
\begin{equation}
  \phi \equiv \frac{N_\mathrm{QSO}}{V_\mathrm{bin}},
\end{equation}
where $N_\mathrm{QSO}$ is the number of quasars with magnitude
$M_\mathrm{min}\leq M<M_\mathrm{max}$ and redshift
$z_\mathrm{min}\leq z<z_\mathrm{max}$, and
\begin{equation}
  V_\mathrm{bin} = \int_{M_\mathrm{min}}^{M_\mathrm{max}}\mathrm{d}M
  \int_{z_\mathrm{min}}^{z_\mathrm{max}}\mathrm{d}z\, f(M, z)\,\frac{\mathrm{d}V}{\mathrm{d}z},
  \label{eqn:vi}
\end{equation}
is the effective volume of the bin. The inclusion of the survey selection function
$f(M,z)$ (Section~\ref{sect:datahom}) in Equation~(\ref{eqn:vi}) accounts for
what are sometimes called ``incomplete bins''
\citep{2006AJ....131.2766R}.  The comoving volume element $\mathrm{d}V/\mathrm{d}z$ is
given by
\begin{equation}
  \frac{\mathrm{d}V}{\mathrm{d}z}=\frac{\mathrm{d}V}{\mathrm{d}z\,\mathrm{d}\Omega}\times A\times\frac{4\pi}{41253},
\end{equation}
where $A$ is the survey area in deg$^2$, and 
\begin{equation}
  \frac{\mathrm{d}V}{\mathrm{d}z\,\mathrm{d}\Omega}=\frac{c}{H_0}\frac{d_L^2\left(z\right)}
       {\left(1+z\right)^2\sqrt{\Omega_\mathrm{m}\left(1+z\right)^3+\Omega_\Lambda}}
  \label{eqn:dvdzdo}
\end{equation}
denotes the comoving volume element per unit solid angle
\citep{1999astro.ph..5116H}. The resulting luminosity function $\phi$ has units of
$\mathrm{cMpc}^{-3}\mathrm{mag}^{-1}$.

We evaluate the double integral in Equation~(\ref{eqn:vi}) by the Euler method,
i.e., by simply summing over the ``tiles'' of the selection function in $M$ and $z$
without interpolation.
This results in $V_i=0$ for a few quasars, which are subsequently removed
from our analysis\footnote{
  We note that the interpolation of sometimes coarse
  selection functions is not straightforward due to their strong gradients.
  The presence of objects with $V_i=0$ implies that the selection function has systematic errors.}.
In each bin we estimate the uncertainty in the luminosity function by
assuming Poisson statistics \citep{1986ApJ...303..336G} for the number
of quasars, i.e.\ assuming negligible uncertainty in the selection function.
While this is a reasonable approximation for small surveys, large surveys
with negligible Poisson errors are instead limited by rarely quantified systematic
errors due to implicit assumptions in their selection functions.
The resultant binned luminosity function estimates
are shown by the circles in Figure~\ref{fig:mosaic}.

From Figure~\ref{fig:mosaic} we see that the distribution of luminosity
function values in each redshift bin are suggestive of a double power
law form for the QLF.  We will fit such a form below. 
However, in several redshift bins we note a suspicious decline of the
luminosity function at the faint limit of several surveys, for example in
the 2SLAQ sample at $z<2.2$, and in the SDSS sample at $z<1.8$ and $z\sim 4$.
The inconsistency between the SDSS faint end and the deeper
2SLAQ QLF indicates that the SDSS selection function is
systematically overestimated at its magnitude limit.
At $z\sim 4$ the SDSS faint end QLF is inconsistent with the fainter
\citet{2011ApJ...728L..26G} QLF. We identify such discrepant bins
by eye, discard the contributing quasars from our analysis, and set their
selection function values to zero.
The discarded magnitude bins are shown in Figure~\ref{fig:mosaic} by
open circles.

%% GW: Maybe say what we mean by one-sigma uncertainty: This is a credible interval from the posterior distribution. Do we use the highest posterior density interval or the equal-tailed interval?
\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{mosaic_small.pdf}
    % mosaic.py 
  \end{center}
  \caption{Homogenised quasar luminosity functions at rest-frame wavelength
    $\lambda=1450$\,\AA\ in redshift bins from $z=0.1$ to $6.5$.
    The symbols show our inferred binned luminosity functions from various
    data sets:
    \citet[red]{2010AJ....139.2360S},
    \citet[green]{2009MNRAS.392...19C},
    \citet[dark blue]{2013ApJ...773...14R},
    \citet[light blue]{2011ApJ...728L..26G},
    \citet[yellow]{2016ApJ...829...33Y},
    \citet[DR7 brown]{2013ApJ...768..105M},
    \citet[Stripe 82 teal]{2013ApJ...768..105M},
    \citet[pink]{2016ApJ...833..222J},
    \citet[orange]{2010AJ....139..906W}, and
    \citet[grey]{2015ApJ...798...28K}.
    Open circles in corresponding colours indicate magnitude bins
    excluded due to incompleteness in the respective data sets.  The
    number of AGN before this selection is shown in parentheses; the
    selected number of AGN is shown outside parentheses.  In each
    redshift bin, the black curve shows our fiducial double power law
    model fit, which is represented by the median of the posterior
    probability distribution function.  The grey shaded area shows the
    one-sigma (68.26\%) uncertainty.  See Sections~\ref{sec:binnedlf}
    and \ref{sec:bins} for further details.}
  \label{fig:mosaic}
\end{figure*}

@@

%%GW: Added a note on discarded redshift bins
%%GW: Check on final Nqso needed! The number at z~5 is inconsistent with Fig. 3
\begin{table*}
  % bins_tabulate.py; Nqso added by hand
  \caption{\textbf{Check on final Nqso needed! The number at $z\sim5$ is inconsistent with Fig. 3. We could also have two columns for the number of quasars: before and after magnitude restrictions.}
    Posterior median double power law luminosity function parameters
    and their $1\sigma$ (68.26\%) statistical errors in
    various redshift bins shown in Figure~\ref{fig:evoln}.
    The luminosity function parameters
    $\phi_*$, $M_*$, $\alpha$, and $\beta$ are defined in
    Equation~(\ref{eqn:dpl}), with $\beta$ denoting the faint-end
    slope.  Quasars in each bin have redshifts $z_\mathrm{min}\leq z <
    z_\mathrm{max}$ with a sample mean $\langle z\rangle$.
    The number of quasars in each bin is given by $N_\mathrm{QSO}$.
    %The corresponding luminosity
    %functions are shown in Figure~\ref{fig:mosaic}.  See
    %Section~\ref{sec:bins} for further details.
    }
  \label{tab:bins}
  \begin{tabular}{lccr....}
    \hline
    $\langle z\rangle$ &
    $z_\mathrm{min}$ &
    $z_\mathrm{max}$ &
    $N_\mathrm{QSO}$ &
    \multicolumn{1}{c}{$\log_{10}(\phi_*/$} &
    \multicolumn{1}{c}{$M_*$} &
    \multicolumn{1}{c}{$\alpha$} &
    \multicolumn{1}{c}{$\beta$} \\
